---
title: "Multi-Classification of Dry Beans Using Machine Learning"
author: "Richard Acquah-Sarpong"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  rmdformats::readthedown:
    highlight: tango
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(message=F,warning=F,echo=T,fig_height=5,fig_width=4, out.extra = "", fig.align= "center")
```

\newpage

```{r, include=FALSE}
## Library
library(knitr)
library(dplyr)
library(ggplot2)
library(viridis)
library(hrbrthemes)
library(corrplot)
library(class)
library(caret)
library(rpart)
library(maptree)
#library(pca3d)
library(MASS)
library(tibble)
library(randomForest)
library(e1071)
library(kernlab)
library(kableExtra)
library(gridExtra)
library(reshape)
library(GGally)
library(readr)
library(readxl)
library(float)

theme_set(theme_light())
```


```{r}
#load data
labeled <- read.csv('labeled.csv') %>% dplyr::select(-X)
sampA <- read.csv('samp.A.csv')%>% dplyr::select(-X)
sampB <- read.csv('samp.B.csv')%>% dplyr::select(-X)
sampC <- read.csv('samp.C.csv')%>% dplyr::select(-X)
#convert Class into factor
labeled$Class <- as.factor(labeled$Class)

#set up a new variable 'Roundness'
#Roundness = 4*Area*pi/(perimeter)^2 (refer to the dry bean paper)
Roundess <- 4*pi*labeled$Area/(labeled$Perimeter)^2
labeled <- add_column(labeled, Roundness = Roundess, .after = 7)
sampA$Roundness <- 4*pi*sampA$Area/(sampA$Perimeter)^2
sampB$Roundness <- 4*pi*sampB$Area/(sampB$Perimeter)^2
sampC$Roundness <- 4*pi*sampC$Area/(sampC$Perimeter)^2

#check for duplicate rows
dup.rows = sum(labeled%>%duplicated(), sampA%>%duplicated(),
               sampB%>%duplicated(),sampC%>%duplicated())
```


## Introduction and Problem Statement

Dry bean- Phaseolus vulgaris L. is a major cultivated grain species in the genus Phaseolus that is widely consumed worldwide for its edible legume and pea pods (Heuze et al., 2015). Nevertheless, **selecting the best seed species is one of the main concerns for both bean producers and the market**. Since different genotypes are cultivated worldwide, it is important to separate the best seed variety from the mixed dry bean population, otherwise the market value of these mixed species of beans could drop enormously (Varankaya & Ceyhan, 2012). For more information on background problem and data, please refer to the associated article "Koklu, M., & Ozkan, I. A. (2020). Multiclass classification of dry beans using computer vision and ma-chine learning techniques. Computers and Electronics in Agriculture, 174, 105507." 


## Aim of the Project
**The aim of this project is to develop a machine learning method to perform a multi-classification of dry beans that could predict the net worth of a given bean species harvested from a 'population cultivation' from a single farm when presented in the market.**


## Data

There are two datasets used in this project; The 'labeled' and 'unlabeled' datasets. The labeled dataset will be used to train the various ML models, hence also refered to as the training dataset. The training dataset contains 3000 observations and 8 variables. The dependent variable has 6 levels (Classes): BOMBAY, CALI, DERMASON, HOROZ, SEKER, and SIRA - These classes represent the different varieties of dry beans. In the training set, each class  has 500 observations. 

The unlabeled dataset is made up of three seperate samples, namely, Sample A, B, and C. these samples have 7 variables (excluding the dependent variable(Class) -thus unlabeled).  The total observations for sample A, B, and C are 777, 1373, and 982 respectively. Roundness, which is the measure of how closely the shape of beans approaches a perfect circle, was calculated and added as an additional predictor variable to both labeled and unlabeled datasets (Koklu & Ozkan, 2020). Tables 1 through 4 show the summary statistics of the variables in the labeled data, Sample A,  B, and C, respectively .


```{r, warning=F, message=F}
summary.stats <- round(as.data.frame((labeled[,-9])%>%psych::describe())%>%dplyr::select(n,mean, sd, median, min, max, range, se), 2)
kbl(summary.stats, caption="Statistical distribution of features of dry beans varieties (in pixels) - Label")%>%kable_classic(full_width = F, html_font = "Cambria")%>%kable_material_dark()
```